Hybrid MILP to efficiently and accuratly \\ solve hard DNN verification instances

### 2024-11-25 ICLR25/Paper_noProof.tex
edit
about propositon 1

\begin{abstract}
$\alpha,\beta$-CROWN has won the last 4 VNNcomp(etitions), as the DNN verifier with the best 
trade-off between accuracy and runtime. VNNcomp however is focusing on relatively easy verification instances. In this paper, we consider {\em harder} verification instances, on which $\alpha,\beta$-Crown displays a large number ($20-58\%$) of undecided instances, that is, instances that can neither be verified, nor an explicit attack can be found. Enabling larger time-outs for $\alpha,\beta$-Crown only improves verification rate by few percents, leaving a large gap of undecided instances while already taking a considerable amount of time. Resorting to other techniques, such as complete verifiers, does not fare better even with very large time-outs: They would theoretically be able to close the gap, but with an untractable runtime on all but small {\em hard} instances.

In this paper, we study the MILP encoding of ReLU-DNNs, and provide new insights in the LP relaxation. This allows us to carefully craft a {\em partial MILP} solution which selects automatically few neurons encoded as integer variables, the rest using the LP relaxation. Compared with previous attempts, we can reduce the number of integer variables by around 4 times while maintaining the same level of accuracy. Implemented in {\em Hybrid MILP}, calling first $\alpha,\beta$-Crown with a short time-out to solve easier instances, and then partial MILP for those for which $\alpha,\beta$-Crown fails, produces a very accurate yet efficient verifier, reducing tremendously the number of undecided instances ($8-15\%$), while keeping a reasonable runtime ($46s-417s$ on average per instance).
\end{abstract}

@@

{\bf Reproducibility Statement:} 
We tested twice outlier results to confirm them, making sure of reproducibility on the given hardware. Precise details on the settings used are provided in the appendix.
Additional results are also provided in the appendix.
Tested DNNs as well as MNIST and CIFAR10 DataSet are freely available.
The source code of Hybrid MILP will be provided on GitHub after acceptance (needing Gurobi as well as $\alpha,\beta$-Crown).

@@

>>>>

\begin{abstract}
$\alpha,\beta$-CROWN has won the last 4 VNNcomp(etitions), as the DNN verifier with the best 
trade-off between accuracy and runtime. VNNcomp however is focusing on relatively easy verification instances. In this paper, we consider {\em harder} verification instances, on which $\alpha,\beta$-Crown displays a large number ($20-58\%$) of undecided instances, that is, instances that can neither be verified, nor an explicit attack can be found. Enabling larger time-outs for $\alpha,\beta$-Crown only improves verification rate by few percents, leaving a large gap of undecided instances while already taking a considerable amount of time. Resorting to slow complete verifiers, does not fare better even with very large time-outs: They would theoretically be able to close the gap, but with an untractable runtime on all but small {\em hard} instances.

In this paper, {\color{blue}
%we study the MILP encoding of ReLU-DNNs.
%, and provide new insights in the LP relaxation. 
we propose a novel Utility function that selects neurons to be encoded with accurate but costly integer variables in a {\em partial MILP} problem. The novelty resides in the use of 
the solution of {\em one} (efficient LP) solver to accurately compute a selection $\varepsilon$-optimal for a given input. }
%This allows us to carefully craft a {\em partial MILP} solution which selects automatically few neurons encoded as integer variables, the rest using the LP relaxation. 
Compared with previous attempts, we can reduce the number of integer variables by around 4 times while maintaining the same level of accuracy. Implemented in {\em Hybrid MILP}, calling first $\alpha,\beta$-Crown with a short time-out to solve easier instances, and then partial MILP for those for which $\alpha,\beta$-Crown fails, produces a very accurate yet efficient verifier, reducing tremendously the number of undecided instances ($8-15\%$), while keeping a reasonable runtime ($46s-417s$ on average per instance).
\end{abstract}

@@

{\bf Reproducibility Statement:} 
We tested twice outlier results to confirm them, making sure of reproducibility on the given hardware. Precise details on the settings used are provided in the appendix.
Additional results {\color{blue}(e.g. ablation studies)} are also provided in the appendix.
Tested DNNs as well as MNIST and CIFAR10 DataSet are freely available.
The source code of Hybrid MILP will be provided on GitHub after acceptance (needing Gurobi as well as $\alpha,\beta$-Crown).

@@

\bigskip

### 2024-11-25 ICLR25/introduction.tex
typo
about propositon 1

Our three main contributions address the challenges of verifying {\em hard} DNNs efficiently:
\begin{enumerate}
	\item  Our first contribution studies the {\em LP relaxation} of the exact MILP encoding of ReLUs. {\color{blue} We establish in Proposition \ref{LP} its equivalence with the so-called "triangular abstraction"}.

	\item {\color{blue} It allows us to choose through a novel Utility function 
	few neurons to encode with the exact MILP encoding, while others are treated with the efficient LP relaxation, giving rise to partial MILP (pMILP). Specifically, the novelty of Utility resides in the use of the solution to an (efficient) full LP call on the node $z$ we want to bound. Utility can then precisely evaluate how much accuracy is gained by switching neuron $a$ from LP (solution of the LP call) to the exact MILP encoding of ReLU (exact computation from the solution, which can be made thanks to Proposition \ref{LP}), with a proved bound on the precision (Proposition \ref{prop2}). pMILP is much more efficient than previous attempts \cite{DivideAndSlide}, necessitating $\approx 4$ times less integer variables for the same accuracy (Table \ref{tab:example1}), because Utility focuses on the solution local to the input, rather than on the DNN. To the best of our knowledge, this is the first time such a solution of an (LP) call is used to evaluate the contribution of each neuron, including heuristics for BaB, e.g. \cite{ferrari2022complete,crown}.}

>>>>

{\color{blue} Our main contributions address the challenges of verifying {\em hard} DNNs efficiently:}
\begin{enumerate}
	%\item  Our first contribution studies the {\em LP relaxation} of the exact MILP encoding of ReLUs. {\color{blue} We establish in Proposition \ref{LP} its equivalence with the so-called "triangular abstraction"}.

	\item {\color{blue} We designed a novel Utility function to choose few neurons to encode with the exact MILP encoding, while others are treated with the efficient LP relaxation, giving rise to partial MILP (pMILP). Specifically, the novelty of Utility resides in the use of the solution to an (efficient LP) solver on the node $z$ we want to bound. Utility can then precisely evaluate how much accuracy is gained by switching neuron $a$ from LP (solution of the LP call) to the exact MILP encoding of ReLU (exact computation from the solution, which can be made thanks to Proposition \ref{LP}), with a proved bound on the precision (Proposition \ref{prop2}). pMILP is much more efficient than previous attempts \cite{DivideAndSlide}, necessitating $\approx 4$ times less integer variables for the same accuracy 
	(Table \ref{tab:example1}), because Utility focuses on the solution local to the input, rather than on the DNN. To the best of our knowledge, this is the first time such a solution of an (LP) call is used to evaluate the contribution of each neuron, including heuristics for BaB, e.g. \cite{ferrari2022complete,crown}.}

### 2024-11-25 ICLR25/valabstraction.tex
about propositon 1

The crucial factor in such an approach is to {\em select} few opened neurons in $X$ which are the most important for the accuracy. An extreme strategy was adopted in \cite{DivideAndSlide}, where only nodes of the immediate previous layer could be opened, and the measure to choose a neuron $a$ to open to compute the values for neuron $b$ was to consider $|W_{ab}| (\UB(a)-\LB(a))$. To obtain a more accurate measure, that can also scale to deeper layers and is not limited to open nodes from the immediate previous layer, we first study the LP relaxation of (\ref{eq11}). We obtain:

@@

{\color{blue}
To the best of our knowledge, we could not find such an equivalence stated in the bibliography, although the proof is simple. This result will be instrumental for our Utility function (Section \ref{sec4}).}

>>>>

The crucial factor in such an approach is to {\em select} few opened neurons in $X$ which are the most important for the accuracy. An extreme strategy was adopted in \cite{DivideAndSlide}, where only nodes of the immediate previous layer could be opened, and the measure to choose a neuron $a$ to open to compute the values for neuron $b$ was to consider $|W_{ab}| (\UB(a)-\LB(a))$. To obtain a more accurate measure, that 
%can also scale to deeper layers and 
is not limited to open nodes from the immediate previous layer, 
{\color{blue} we invoke a folklore result on the LP relaxation of (\ref{eq11}), for which we provide a direct and explicit proof:}

@@

%{\color{blue}
%To the best of our knowledge, we could not find such an equivalence stated in the bibliography, although the proof is simple. This result will be instrumental for our Utility function (Section \ref{sec4}).}
